---
title: "REAVER extraction"
author: "Andreas Schönberg"
date: "26 September 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Description

The function REAVER extraction is used to extract values from multilayer rasters for several spatial polygons. Further it calculates the standrt devation, sum, mean, maximum and minimun values for each polygon from the pixel-based vales. Returns a data frame. 
NOTE: requires a RSAGA environment!

Development information:
The function is much faster than the standart raster extraction fucntion. It rasterizes the polygons, extracts the values from the input multilayer raster to the rasterized poylgons and creates a dataframe with the values from each raster within the stack or brick and for each pixel within the polygons. Further the function calculates the sd, sum, mean, min and max for each raster in each polygon.

With rasterizing the poylgons, informations in the spatial polygons dataframe are lost. The function requires a ID with unique numeric values for each polygon. If "set id = TRUE" a column with an id from 0 to length of polygons will be created. If"set id=FALSE" and an column "ID" exist the ID will be used.
The ID will be used for Rownames of the dataframe to identifiy the pinput polygons

### Output data
The returning dataframe has the polgon id in rownames and the input rasters from the multilayer in colums for sd,sum,mean,min and max (5 colums for each raster in the multilayer).

### Outlook
Next update will include optinal selection of mathematic parameters and more parameters.

Requires "plyr" package
```{r cars}
#' Mandatory: Reaver Extraction
#' 
#' @description Optional: Function to extract values from multilayer objects for several spatial polygons
#' and calculate multiple statistic values with one value per polygone. returns a single data frame.
#' @name Mandatory Reaver 
#' @export Mandatory Reaver

#' @param Mandatory if function: poly - a shp with polygones
#' @param Mandatory if function: multilayer - a rasterbrick, needs same CRS and extensions like the shp
#' @param Mandatory if function: set_ID - bolean, if TRUE an ID is added to the polygones, default is TRUE
#' @param Mandatory if function: stats - bolean, if TRUE returns a data frame with sd,mean,sum,min and max values for each polygone, default is TRUE
#' @param Mandatory if function: spell - bolean, if TRUE the function will call the workflow position, default is TRUE
#' @param Mandatory if function: name - character in "" - paste a name to the Object ID, default is "" empty.

#note: v1.1 now paste a name to the object ID, renewed structur

Reaver_extraction <- function(poly,multilayer,set_ID=TRUE,name="") {
  
  #infun Test###
  #poly=poly
  #multilayer=stk
  ###############
  
 
  cat(" ",sep = "\n")
  cat("### Reaver starts Extraction of Values ###")
  cat(" ",sep = "\n")

  #add ID column
  if (set_ID==TRUE){
    poly@data$ID=seq(1,length(poly),1)
  }else{
    poly=poly  } #if input data has a defined "ID" column
  
  #rasterize polygones , mask the multilayer to rasterized polygones and add the rasterized polygone 
  cat(" ",sep = "\n")
  cat("### Reaver rasterizes the polygones for",nlayers(multilayer),"Layers ###")
  rastrize <- raster::rasterize(poly,multilayer,field="ID")
  masked <- raster::mask(multilayer,rastrize)
  masked_ID <- raster::addLayer(masked,rastrize)
  
  #extract Values , handle NA and get df
  cat(" ",sep = "\n")
  cat("### Reaver extracts the Values from ",nlayers(multilayer),"Layers ###")
  getval <- raster::getValues(masked_ID)
  clean <- na.omit(getval)
  df_clean <- as.data.frame(clean)
  
  ### statistic path
  # calculate statistic, types can be outcommended
  cat(" ",sep = "\n")
  cat("### Reaver calculates statistic values ###")
  sd   <-ddply(df_clean,.(layer),colwise(sd))
  mean <-ddply(df_clean,.(layer),colwise(mean))
  sum <-ddply(df_clean,.(layer),colwise(sum))
  min <-ddply(df_clean,.(layer),colwise(min))
  max <-ddply(df_clean,.(layer),colwise(max))
    
  # rename for each typ and merg df
  colnames(sd)[2:length(sd)]     <- paste(colnames(sd)[2:length(sd)],"sd",sep = "_")
  colnames(mean)[2:length(mean)] <- paste(colnames(mean)[2:length(mean)],"mean",sep = "_")
  colnames(sum)[2:length(sum)]   <- paste(colnames(sum)[2:length(sum)],"sum",sep = "_")
  colnames(min)[2:length(min)]   <- paste(colnames(min)[2:length(min)],"min",sep = "_")
  colnames(max)[2:length(max)]   <- paste(colnames(max)[2:length(max)],"max",sep = "_")
   
    
  df <- merge(sd,mean,by="layer")
  df <- merge(df,sum,by="layer")
  df <- merge(df,min,by="layer")
  df <- merge(df,max,by="layer")
  #row.names(df) <-paste0(name,row.names(df),sep="")
  row.names(df) <-paste0(name,poly$ID)
  #df
  cat(" ",sep = "\n")
  cat(" ",sep = "\n")
  cat("### Reaver has finished ###")
  cat(" ",sep = "\n")
  return(df)
    
} 
```


```{r pressure, echo=FALSE}

```

